---
title: "p8105_hw1_ll3344"
author: "Lusha Liang"
date: "9/13/2020"
output: github_document
---

# Problem 1

## Creating a data frame

Here we are creating a data frame comprised of:

* A random sample of size 10 from a standard normal distribution
* A logical vector indicating whether elements of the sameple are greater than 0
* A character vector of length 10
* A factor vector of length 10, with 3 different factor "levels"

```{r dataframe}
#load tidyverse
library(tidyverse)

#create our vectors
set.seed(123)
vec_normal = rnorm(10)

vec_logical = as.logical(vec_normal > 0)

vec_char = c("Learning", "R", "is", "so", "much", "fun", "indeed", "I", "do", "declare")

vec_factor = factor(c("alpacas", "sloths", "pandas", "pandas", "sloths", "alpacas", "alpacas", "alpacas", "sloths", "pandas"))

# create our data frame using the above defined vectors                    
df_prob_1 = tibble(
  vec_normal,
  vec_logical,
  vec_char,
  vec_factor
)

df_prob_1

```

## (Attempting) means
The mean of vec_normal is `r mean(vec_normal)`. 
The mean of vec_logical is `r mean(vec_logical)`. 
The mean of vec_char is `r mean(vec_char)`. This returns an error because the argument is not numeric or logical and therefore taking the mean does not make sense. 
The mean of vec_factor is `r mean(vec_factor)`. This also returns an error because the argument is not numeric or logical. 

We can convert variables from one type to another in certain cases. Below is code showing our attempt to convert logical, character, and factor variables to numeric. 

```{r numeric, results = "hide", warning = FALSE}
as.numeric(vec_logical)
as.numeric(vec_char)
as.numeric(vec_factor)
```

In the first case we will be successful in converting vec_logical to numeric in that R assumes that FALSE translates to 0 and TRUE translates to 1 in numeric form. Since we have 5 "TRUE"s and 5 "FALSE"s, this explains why our mean was 0.5. 

In the case of the character and factor vectors, R will not be successful in converting these vectors to numeric because there are no built in assumptions as to what number (for example) alpacas should translate to. This is why R returned an error when we tried to take the means of these vectors. 

## Additional data manipulation
Finally, we will do the following data manipulation: 

1. Convert the logical vector to numeric, and multiply the random sample by the result
2. Convert the logical vector to a factor, and multiply the random sample by the result
3. Convert the logical vector to a factor and then convert the result to numeric, and multiple the random sample by the result

```{r manipulation}
# Step 1
logic_1 = as.numeric(vec_logical)
logic_1 * vec_normal

# Step 2
logic_2 = as.factor(vec_logical)
logic_2 * vec_normal

# Step 3
logic_3 = as.numeric(logic_2)
logic_3 * vec_normal
```

As shown above R returns an error when we try to multiply a factor in the second step, but we are able to convert the factor to numeric, then successfully multiply. Notably we see that R assumes factor levels start at the number 1 so that the first level is coded as 1 and the second level as 2, whereas FALSE and TRUE correspond to 0 and 1, respectively. 


